---
title: "API + Maps: Spatial Interpolation"
author: "FCA Collin, Ph.D - <https://FCACollin.github.io/Latarnia>"
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
bibliography: ref/reference/bibtex/ref.bib
output_format:
    html_document:
        toc: TRUE
        toc_float: FALSE
        code_folding: 'show'
        theme: 'lumen'
        number_sections: TRUE

---

Introduction
============

The low quality of air in Poland is responsible for the anticipated death
of 50000 people every year. Not only fatalities, life comfort of inhabitants
is impacted with smog triggering chronicle respiratory
difficulties, nauseas or headaches. Fight against smog encompasses a large range
of action supported by public funds, from promotion of more recent furnace to
the local ban of coal combustions.

A cornerstone of public measure lays in the ability to describe an initial
state and monitor its evolution; it makes possible the investigation of the
expected improvement associated with a strategy. In this
respect, a  network of station measuring air quality across Poland provides
hourly evaluation of various pollutant. The station data are either directly
delivered through synthetic maps or less refined as raw data (a measure at a
give date and time for a specific pollutant monitored by a peculiar station).

Stations provide air quality estimation in a specific location and provide
therefore a localised description of the air quality. The number of station
changes with time, with a general trend to increase the number of stations
available. In this condition, the stated of air quality is not available
for the whole territory.

To account for spacial discrete distribution of station, and
variation with time of data acquisition availability the complete network
of air quality monitoring can be interpolated so as to obtain a local estimation
in any position within the map. This will help representing the general
situation of air quality in Poland and also understanding the limit of the current
air quality monitoring schemes. The approach presented hereafter is a possible
general framework for the generation of interpolated maps along with
prediction variance estimations.

```{r setupKnitr, echo = FALSE, message = FALSE, warning = FALSE}
#[FC 191010 08:15]# {{{

fig.height <- 5;
fig.width <- 16/9 * fig.height;
knitr::opts_chunk$set(
  warning    = FALSE,
  message    = FALSE,
  echo       = TRUE,
  results    = 'hide',
  cache      = FALSE,
  fig.width  = fig.width,
  fig.height = fig.height,
  dpi        = 300,
  dev.args = list(bg = 'transparent'),
  dev = c('png', 'pdf')
  );

knitr::read_chunk("R/PM10.R")

###}}} -- vim fold
```


Material and methods
====================

Principle
---------

The general idea is to retrieve and select air quality data from station
scattered across the country and represent them on a map. The stations
represent singular points, the scattered values are used for interpolation
so as to estimated the air quality over all territory in a *raster*.
The raster, the country boundaries (*shape*), the station measure (points)
are then stacked over an open street map background.


Data
----

Updated PM10 data are available via an API interface:
<https://powietrze.gios.gov.pl/pjp/content/api>.

The map shape was downloaded from the 
Geographic Information System of the COmmission (GISCO), a service
within Eurostats. The GISCO proposes a nomenclature of territorial units for
Statistics, named NUTS. The NUTS of the year 2016 was downloaded before running
the application[^1].

Packages
--------

- The API queries were managed with `httr` [@Wickham2019]
  and `jsonlite` [@Ooms2014].
- The geographic data were handled with `sp` [@Pebesma2005;@Bivand2013],
  `rgdal` [@Bivand2019], `raster` [@Hijmans2020] and `gstat` [@Graeler2016].
- The interactive rendering was managed with `leaflet` and `htmlwidgets`.


```{r settings}
```


Results
=======


PM10 data
---------

The API interface was used to query the list of available stations.

```{r data_station}
```

```{r table_station, results = 'asis', echo = FALSE}
```

Then, for every station, the list of sensors were stacked in a table.

```{r data_sensor}
```

```{r table_sensor, results = 'asis', echo = FALSE}
```

Finally, measures are extracted for a selected air quality parameter,
(e.g. PM10).

```{r data_measure}
```

```{r table_measure, results = 'asis', echo = FALSE}
```


GIS objects
-----------

The shape file was previously downloaded[^1], loaded into the session and
restricted to the voivodeships (second administrative division) in
Poland. The shape Coordinate Reference System was transformed in
the World Geodetic System 1984 (WGS84)

```{r data_geo_shape_pl}
```

```{r map_geo_shape_pl, fig.cap="Poland shape file."}
```

The PM10 data prepared earlier were restricted to the latest
date and time with recorded by a sufficient number of station.
The points were then added to the available shape previously obtained.

```{r data_geo_points}
```

```{r map_geo_points, fig.cap="Addition of station location to the shape file."}
```

Density of points was heterogeneous, while some area were poorly monitored,
some other area cumulated locally many stations.
The map was then divided into a grid composed of 300 rows and 300 columns,
defining therefore 90000 squares; the grid and squares, are later referred as
raster and pixels. The value of each pixel was
estimated as the average of the local value falling in the area covered by that
pixel.


```{r data_gis_raster}
```

```{r graph_gis_raster, fig.cap="From station to pixel value estimation."}
```

A linear interpolation was computed to fill the raster gap.

```{r data_gis_raster_intpl}
```

```{r graph_gis_raster_intpl, fig.cap="Linear interpolation of local PM10 measures."}
```

Eventually, zone contour were delimited and a more appropriate colour scale
was proposed.

```{r graph_gis_raster_intpl_2, fig.cap="Automatic detection of contours."}
```


Interactive map
---------------

Finally, the raster, the shape and the point were overlayed on a Open Street
Map background. The `leaflet` package proposes a convenient high-level graphical
programming interface were layers are added to each other. Note that a
colour scale was refined so as to correspond to Polish standards for which
a green-to-red gradient is limited to PM10 concentrations between 0 and 150.


```{r data_map_leaflet}
```

<iframe
src="html/html_widget/PM10_map.html" width="100%" height="600" frameborder="0"
style="border:0;" allowfullscreen=""></iframe>

<a href="html/html_widget/PM10_map.html" target="_blank">Open in a new tab.</a>


Session information
-------------------

```{r sessionInfo, echo = FALSE, results = "markup"}
sessionInfo()
```


Discussion
==========

The Polish GOŚ API is a easy-to-use interface to query the recent measure
accounting for air quality at a national scale. The projection of this
this information to a map layer and linear interpolation is rather
straightforward. The high-level function leaflet is of particular interest
to combine the different map layers into a fashionable interactive map
easily integrated into a web page as a html widget.

If the map can be updated, the information projected is nonetheless static.
It could be of interest to add a reactive component such as proposed by the
`Shiny` package so as to be able to choose a possible date.
What is more, an animation cycling through available measure date and time
would help perceiving the temporal trend.



References
==========


[^1]: <https://ec.europa.eu/eurostat/cache/GISCO/distribution/v1/nuts-2016.html>

[modeline]: # ( vim: set foldlevel=0 spell spelllang=en_gb: )

